---
title: "Data Transformation"
author: 
  - "Travis Hinkelman"
date: "2020-12-04"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#699CC6",
  secondary_color = "#F9A134",
  inverse_header_color = "#FFFFFF",
  extra_css = list(".title-slide" = list("background-color" = "#FFFFFF"),
                   ".title-slide, .title-slide h1, .title-slide h2, .title-slide h3" = 
                     list("color" = "#699CC6")))
```

## Texas Housing Dataset

Texas housing dataset is included in `ggplot2` package (included in `tidyverse`). 

```{r echo = TRUE, message=FALSE}
library(tidyverse)
```

--

Open documentation to read dataset description.

```{r echo = TRUE, eval=FALSE}
?txhousing
```

--

Get dimensions of dataframe.

```{r echo = TRUE}
dim(txhousing)
```

--

Or just get the number of rows.

```{r echo = TRUE}
nrow(txhousing)
```

---

## Texas Housing Dataset

Get the column names.

```{r echo = TRUE}
names(txhousing)
```

--

Get first five rows. 

```{r echo = TRUE}
head(txhousing, 5)
```

---

## Texas Housing Dataset

Get rows 101-110.

```{r echo = TRUE}
txhousing[101:110, ]
```

---

## Filter

Find all unique values in the year column and print the first 10.

```{r echo = TRUE}
unique(txhousing$year)[1:10]
```

--

Filter `txhousing` to keep rows where `year` is 2008. 

```{r echo = TRUE}
head(filter(txhousing, year == 2008))
```

---

## Filter

Filter `txhousing` to keep rows where `year` is 2008 **and** `city` is Lubbock. 

```{r echo = TRUE}
filter(txhousing, year == 2008 & city == "Lubbock")
```

---

## Filter

Filter `txhousing` to keep rows where `year` is 2008 **or** `city` is Lubbock. 

```{r echo = TRUE}
filter(txhousing, year == 2008 | city == "Lubbock")
```

---

## Filter

Filter `txhousing` to keep rows where `year` is 2006-2010. 

```{r echo = TRUE}
df1 <- filter(txhousing, year > 2005 & year <= 2010)
unique(df1$year)
```

--

`%in%` provides more convenient syntax to get same result.

```{r echo = TRUE}
df2 <- filter(txhousing, year %in% 2006:2010)
unique(df2$year)
```

---

## Filter

Filter `txhousing` to keep rows where `city` is not equal to Lubbock. 

```{r echo = TRUE}
df3 <- filter(txhousing, city != "Lubbock")
```

Check if Lubbock is found in `city` column of `df3` dataframe.

```{r echo = TRUE}
"Lubbock" %in% df3$city
```

--

Filter `txhousing` to keep rows where `year` is not in the vector `2006:2010`. 

```{r echo = TRUE}
df4 <- filter(txhousing, !(year %in% 2006:2010))
unique(df4$year)
```

---

## Filter

Filter `txhousing` to keep rows where `median` is missing. 

```{r echo = TRUE}
filter(txhousing, is.na(median))
```

---

## Filter

Filter only retains rows where the expression is `TRUE`, i.e., `NA` is dropped.

```{r echo = TRUE}
nrow(filter(txhousing, median > 0))
```

--

Use `!is.na` to be more explicit about your intent to drop rows with `NA`.

```{r echo = TRUE}
nrow(filter(txhousing, !is.na(median)))
```

--

If you want to retain `NA`, you need to include it in `filter` expression.

```{r echo = TRUE}
nrow(filter(txhousing, is.na(median) | median > 0))
```

---

## Arrange

Arrange `txhousing` to sort ascending by `median` column.

```{r echo = TRUE}
head(arrange(txhousing, median), 3)
```

--

Arrange `txhousing` to sort descending by `median` column.

```{r echo = TRUE}
head(arrange(txhousing, desc(median)), 3)
```

---

## Arrange

If you arrange rows based on multiple columns, each additional column sorts within groupings (if any) defined by previous columns (listed left to right).

```{r echo = TRUE}
arrange(txhousing, desc(city), median)
```

---

## Arrange

Missing values are placed at end of dataframe.

```{r echo = TRUE}
df5 <- tibble(Col1 = c(2, NA, 1), Col2 = c("A", "T", "R"))
arrange(df5, Col1)
```

--

```{r echo = TRUE}
arrange(df5, desc(Col1))
```

---

## Select

Select `date`, `city`, and `median` columns from `txhousing`.

```{r echo = TRUE}
select(txhousing, date, city, median)
```

---

## Select

Drop `city`, `year`, and `month` columns from `txhousing`.

```{r echo = TRUE}
head(select(txhousing, -city, -year, -month), 3)
```

--

```{r echo = TRUE}
head(select(txhousing, -c(city, year, month)), 3)
```

---

## Select

Select consecutive columns.

```{r echo = TRUE}
head(select(txhousing, sales:date), 3)
```

--

Drop consecutive columns (note the parentheses around `sales:date`).

```{r echo = TRUE}
head(select(txhousing, -(sales:date)), 3)
```

---

## Select

A few helper functions available for more flexible `select` operations, e.g., `starts_with`, `ends_with`, `contains`.

```{r echo = TRUE}
head(select(txhousing, contains("e")))
```

---

## Select

Rename while selecting.

```{r echo = TRUE}
select(txhousing, city, year, month, median_sale_price = median)
```

---

## Rename

Rename without selecting.

```{r echo = TRUE}
rename(txhousing, median_sale_price = median)
```

---

## Pipe Operator

A series of operations on a dataframe can create a lot of intermediate outputs.

```{r echo = TRUE}
df7 <- filter(txhousing, city == "Lubbock")
df8 <- arrange(df7, desc(year))
df9 <- select(df8, city, year, month, median)
```

The pipe operator `%>%` takes the results of the function on the left and passes it as the first argument to the function on the right. [Insert `%>%` with keyboard shortcut `Ctrl` + `Shift` + `m`.]

```{r echo = TRUE, eval=FALSE}
txhousing %>% 
  filter(city == "Lubbock") %>% 
  arrange(desc(year)) %>% 
  select(city, year, month, median)
```

---

## Mutate

Create a tibble, sum `Juvenile` and `Adult` columns to create `Total`, divide `Adult` by total to create `PropTotal`, and add `Site` with `A` repeated for every row.

```{r echo = TRUE}
tibble(Day = 1:5, 
       Juvenile = round(runif(length(Day), 0, 100)), 
       Adult = round(runif(length(Day), 0, 20))) %>% 
  mutate(Total = Juvenile + Adult, 
         PropAdult = Adult/Total, 
         Site = "A")
```

---

## Group By and Summarize

```{r echo = TRUE}
txhousing %>% 
  filter(city == "Dallas" & year > 2004) %>% 
  group_by(year) %>% 
  summarize(avg_median = mean(median, na.rm = TRUE)) %>% 
  head(8)
```

---

## Group By and Summarize

```{r echo = TRUE}
txhousing %>% 
  filter(city %in% c("Dallas", "Houston") & year %in% 2006:2010) %>% 
  group_by(city, year) %>% 
  summarize(avg_median = mean(median, na.rm = TRUE))
```

---

## Group By and Filter

```{r echo = TRUE}
txhousing %>% 
  group_by(city) %>% 
  filter(median > quantile(median, probs = 0.9, na.rm = TRUE))
```

---

## Group By and Mutate

```{r echo = TRUE}
txhousing %>% 
  group_by(year) %>% 
  mutate(prop_max = median/max(median, na.rm = TRUE)) %>% 
  select(city, year, month, median, prop_max)
```
