---
title: "ggplot Basics"
author: 
  - "Taylor Spaulding"
date: "2021-01-08"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#699CC6",
  secondary_color = "#F9A134",
  inverse_header_color = "#FFFFFF",
  extra_css = list(".title-slide" = list("background-color" = "#FFFFFF"),
                   ".title-slide, .title-slide h1, .title-slide h2, .title-slide h3" = 
                     list("color" = "#699CC6")))
```

## Graphics Language

1. ggplot has a particular syntax with 7 unique components
  - Data
  - Geometries
  - Transformations
  - Position
  - Scale
  - Coordinate System
  - Grouping
  
2. First 5 components are grouped into layers
3. Layers are combined to produce plots

---

## Pivot Wider

```{r echo = TRUE, out.height = '320px'}
txhousing %>% 
  filter(year %in% 2004:2012 & city == "San Antonio" & 
           month %in% 4:10) %>% 
  pivot_wider(id_cols = year, names_from = "month", 
              values_from = "median")
```

---

## Pivot Wider

```{r echo = TRUE}
det <- tibble(FishID = c(rep("X1", 3), rep("X2", 4), rep("X3", 4)),
              Day = c(1, 3, 4, 1, 2, 3, 4, 1, 5, 6, 7),
              Receiver = c("A", "C", "D", "A", "D", "E", "F", 
                           "A", "B", "C", "E"))
det
```

---

## Pivot Wider

```{r echo = TRUE}
det %>% 
  mutate(Presence = 1) %>% 
  pivot_wider(id_cols = FishID, names_from = "Receiver", 
              values_from = "Presence", values_fill = 0)
```

---

## Pivot Wider

```{r echo = TRUE}
det %>% 
  mutate(Presence = 1) %>% 
  arrange(Receiver) %>% 
  pivot_wider(id_cols = FishID, names_from = "Receiver", 
              values_from = "Presence", values_fill = 0)
```

---

## Pivot Longer

```{r echo = TRUE}
det %>% 
  mutate(Presence = 1) %>% 
  arrange(Receiver) %>% 
  pivot_wider(id_cols = FishID, names_from = "Receiver", 
              values_from = "Presence", values_fill = 0) %>% 
  pivot_longer(cols = -FishID, names_to = "Receiver", 
               values_to = "Presence") %>% 
  head()
```

---

## Pivot Longer

```{r echo = TRUE}
det %>% 
  mutate(Presence = 1) %>% 
  arrange(Receiver) %>% 
  pivot_wider(id_cols = c(FishID, Day), names_from = "Receiver", 
              values_from = "Presence", values_fill = 0) %>% 
  pivot_longer(cols = -c(FishID, Day), names_to = "Receiver", 
               values_to = "Presence") %>% 
  head()
```

---

## Left Join

.pull-left[
```{r echo = TRUE}
det
```
]

.pull-right[
```{r echo = TRUE}
rec <- 
  tibble(Receiver = LETTERS[1:6],
         RKM = c(33, 28, 20, 15, 9, 1))
rec
```
]

---

## Left Join

```{r echo = TRUE}
left_join(det, rec)
```

---

## Left Join

```{r waterfall, echo = TRUE, out.height = '320px'}
left_join(det, rec) %>% 
  ggplot(aes(x = Day, y = RKM, col = FishID)) +
  geom_point(size = 3) +
  geom_line(size = 1)
```

---

## Separate

```{r echo = TRUE}
tibble(Species_Lifestage = c("Chinook_Spawning", "Chinook_Fry",
                             "Steelhead_Spawning", "Steelhead_Fry"),
       Count = c(12, 980, 17, 1234)) %>% 
  separate(col = "Species_Lifestage", 
           into = c("Species", "Lifestage"))
```

---

## Unite

```{r echo = TRUE}
tibble(Species_Lifestage = c("Chinook_Spawning", "Chinook_Fry",
                             "Steelhead_Spawning", "Steelhead_Fry"),
       Count = c(12, 980, 17, 1234)) %>% 
  separate(col = "Species_Lifestage", 
           into = c("Species", "Lifestage")) %>% 
  unite(Species_Lifestage, Species, Lifestage)
```


---

## Unite

```{r echo = TRUE}
tibble(Species_Lifestage = c("Chinook_Spawning", "Chinook_Fry",
                             "Steelhead_Spawning", "Steelhead_Fry"),
       Count = c(12, 980, 17, 1234)) %>% 
  separate(col = "Species_Lifestage", 
           into = c("Species", "Lifestage")) %>% 
  unite(Species_Lifestage, Species, Lifestage, sep = ".")
```

