---
title: "ggplot Continued"
author: 
  - "Taylor Spaulding"
date: "2021-01-22"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#699CC6",
  secondary_color = "#F9A134",
  inverse_header_color = "#FFFFFF",
  extra_css = list(".title-slide" = list("background-color" = "#FFFFFF"),
                   ".title-slide, .title-slide h1, .title-slide h2, .title-slide h3" = 
                     list("color" = "#699CC6")))
```

## ggplot Continued

Topics for Today
  - Grouping
  - Statistics
  - Theming

---

## Load Up Your Packages

```{r echo = TRUE}
library(tidyverse)
```

---

## Load In Some Data

Load in the data from a dataset stored in the Project. 

```{r echo = TRUE}
poke <- read.csv(file.path("~","R","R-Intro-Training","Data", "Pokemon.csv"))

```

---
## Basic Mutations

Make all the factorized fields into characters

Filter the dataset for only Grass, Water, and Fire Types

```{r echo = TRUE}
poke <- poke %>%
  mutate_if(., is.factor, as.character)

poke_mod <- poke %>%
  filter(., stringr::str_detect("(Grass|Water|Fire)", Type.1)) #filter for just those with these types

```

---

## Examine the Data

```{r echo = TRUE}
str(poke_mod)
```

---

## Creating A Simple Plot

```{r echo = TRUE, out.height = '400px', out.width= '500px'}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point()
```

---

## Faceting

Sometimes plots are difficult to read when too many groups are present, faceting can be used to reduce clutter.

`facet_grid([rows] ~ [columns])` is added to our original plot.

A period is used to "ignore" the rows here
```[r]
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  facet_grid(. ~ Type.1)
```

---

## Faceting

```{r echo == TRUE, out.height = '400px', out.width= '500px'}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  facet_grid(. ~ Type.1)
```

---

## Faceting
By changing which side of the "~" we place the variable on we change the orientation.
```{r echo = TRUE, out.height = '300px', out.width= '400px'}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  facet_grid(Type.1 ~ .)
```

---

## Making things Tricky

Lets change the data to make Water Type 6x more abundant
  
```{r echo = TRUE}
poke_mod %>%
  base::split(., .$Type.1) %>% # Split the dataset by type
  purrr::map(~ dplyr::filter(., Type.1 == "Water")) %>% # Filter the dataset for only water pokemon
  base::Filter(function(x) dim(x)[1] > 1, .) %>%
  base::rep(., 5) %>% # Repeat the entire data set 5 times for 6 records
  base::do.call(rbind, .) %>% #bind all the rows together into a single table
  base::rbind(., poke_mod) -> six_times_water # Bind the new table to the old

```
---

## Histograms
.pull-left[  
```{r echo = TRUE, out.height = '300px', out.width= '400px'}
ggplot(data = six_times_water, mapping = aes(x = Attack, fill = Type.1)) +
  geom_histogram() +
  facet_grid(Type.1 ~ .)
```
]

.pull-right[ 
It's hard to read Fire and Grass, because the y-axis scale is determined by Water
]

---

## Facet Axis Scales

We can adjust scales for the x or y axes

`scales = "free_y"`

.pull-left[
```[r]
ggplot(data = six_times_water, 
       mapping = aes(x = Attack, 
       fill = Type.1)) +
  geom_histogram() +
  facet_grid(Type.1 ~ . , 
             scales = "free_y")
```
]

.pull-right[
```{r echo = FALSE, out.height = '300px', out.width= '400px'}
ggplot(data = six_times_water, mapping = aes(x = Attack, fill = Type.1)) +
  geom_histogram() +
  facet_grid(Type.1 ~ . , scales = "free_y")
```
]

---
## Facetting

can include facets for both rows and columns

.pull-left[
Type ~ Legendary
```{r echo = FALSE, out.height = '300px', out.width= '400px', message=FALSE}
ggplot(data = poke_mod, mapping = aes(x = Attack, fill = Type.1)) +
  geom_histogram() +
  facet_grid(Type.1 ~ Legendary,)
```
]

.pull-right[
Legendary ~ Type
```{r echo = FALSE, out.height = '300px', out.width= '400px', message=FALSE}
ggplot(data = poke_mod, mapping = aes(x = Attack, fill = Type.1)) +
  geom_histogram() +
  facet_grid(Legendary ~ Type.1,)
```
]
---

## New Statistics/Summaries

Smoothers turn data (points) into smoothed curves
```
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, col = Type.1)) +
  geom_point() +
  geom_smooth(se = FALSE, span = 0.7)
```
se = FALSE turns off the standard error ribbon

span = 0.7 determines the "wiggliness" of the curve try other combos, higher = less squiggle

---

## Statistics/Summaries

Smoothers turn data (points) into smoothed curves
```{r, out.height = '300px', out.width= '400px'}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, col = Type.1)) +
  geom_point() +
  geom_smooth(se = FALSE, span = 0.7)
```
---

## Statistics and Summaries

`geom_histogram` is another stat we've already used


```{r echo = TRUE, out.height= 300, out.width= 400}
ggplot(data = poke_mod, mapping = aes(x = Attack)) +
  geom_histogram()
```

---

## Statistics and Summaries

We can also make a histogram using `geom_bar` and `stat = "bin"`
geom_histogram is called a "wrapper"

```{r, echo = TRUE, out.height=300, out,width = 400}
ggplot(data = poke_mod, mapping = aes(x = Attack)) +
  geom_bar(stat = "bin")
```

---

## Statistics

Sometimes we want to plot "standard" distributions the `stat_function` geometry is used for this.

Create new x-values for distribution
```{r, echo = TRUE, out.height= 400, out.width=400}
x <- data.frame(x = seq(-4, 4, length = 1000))
```

---
## Statistics

Sometimes we want to plot "standard" distributions the `stat_function` geometry is used for this.

Plot the normal distribution with ggplot
```{r, echo = TRUE, out.height= 300, out.width=400}
ggplot(data = x, mapping = aes(x = x)) +
  stat_function(fun = dnorm)
```

---
## Statistics

Sometimes we want to plot "standard" distributions the `stat_function` geometry is used for this.

Plot the t-distribution with ggplot
```{r, echo = TRUE, out.height= 300, out.width=400}
ggplot(data = x, mapping = aes(x = x)) +
  stat_function(fun = dt, args = list(df = 20))
```

---
## Statistics

Using these statistical functions to extract y-values for x-values in the t-distribution, with 20 degrees of freedom
```{r, echo = TRUE, out.height= 400, out.width=400}
data <- data.frame(x = x, y = dt(x$x, df = 20))
```

---

## Statistics

Get two-tailed critical value for t-distribution with 20 df and alpha = 0.05
```{r, echo = TRUE, out.height= 400, out.width=400}
critical_value <- qt(0.025, 20)
```

---
## Statistics

Make a beautiful graph of the t-distribution
```
ggplot(data = x, mapping = aes(x = x)) +
  stat_function(fun = dt, args = list(df = 20))+
  geom_area(data = subset(data, x <= critical_value), 
            mapping = aes(x = x, y = y), fill = "red") +
  geom_area(data = subset(data, x >= abs(critical_value)), 
            mapping = aes(x = x, y = y), fill = "red") +
  geom_vline(xintercept = c(critical_value, abs(critical_value))) +
  annotate("text", x = -3.5, y = 0.375, 
           label = "alpha = 0.05") +
  annotate("text", x = -3.25, y = 0.05, 
           label = paste("critical value = ", round(critical_value, digits = 3)), col = "red") +
  annotate("text", x = 3.25, y = 0.05, 
           label = paste("critical value = ", round(abs(critical_value), digits = 3)), col = "red")

```
notice use of `stat_function`, `geom_area`, `geom_vline`, and `annotate`
---
## Statistics
Make a beautiful graph of the t-distribution

```{r, echo = FALSE, out.height= 400, out.width=500}
ggplot(data = x, mapping = aes(x = x)) +
  stat_function(fun = dt, args = list(df = 20))+
  geom_area(data = subset(data, x <= critical_value), 
            mapping = aes(x = x, y = y), fill = "red") +
  geom_area(data = subset(data, x >= abs(critical_value)), 
            mapping = aes(x = x, y = y), fill = "red") +
  geom_vline(xintercept = c(critical_value, abs(critical_value))) +
  annotate("text", x = -3.5, y = 0.375, 
           label = "alpha = 0.05") +
  annotate("text", x = -3.25, y = 0.05, 
           label = paste("critical value = ", round(critical_value, digits = 3)), col = "red") +
  annotate("text", x = 3.25, y = 0.05, 
           label = paste("critical value = ", round(abs(critical_value), digits = 3)), col = "red")

```

notice use of `stat_function`, `geom_area`, `geom_vline`, and `annotate`
---

## Statistics

Summarizing y-values at an x-value
.pull_left[
```[r]
ggplot(data = poke_mod, mapping = aes(x = Type.1, y = Speed)) +
  geom_jitter(width = 0.1) +
  stat_summary(geom = "point", fun.y = "median", size = 5, aes(colour = Type.1))
```
]
.pull_right[
`geom_jitter` creates point geometries which are separated randomly

the `stat_summary` call finds the median y-values and plots a point

]
---
## Statistics

Summarizing y-values at an x-value
.pull_left[
```{r echo = FALSE, out.height= 300, out.width=400}
ggplot(data = poke_mod, mapping = aes(x = Type.1, y = Speed)) +
  geom_jitter(width = 0.1) +
  stat_summary(geom = "point", fun.y = "median", size = 5, aes(colour = Type.1))
```
]
.pull_right[
`geom_jitter` creates point geometries which are separated randomly

the `stat_summary` call finds the median y-values

]
---

## Coordinates and Scales

We've already seen how we can assign x and y coordinates, but we can also use the coordinate system to display our plots better

Lets reorder the count data by type for all pokemon types
```{r, out.height= 400, out.width=500}
highest_to_lowest <- poke %>%
  dplyr::group_by(., Type.1) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::arrange(., n) %>%
  dplyr::pull(., Type.1) %>%
  base::unique(.)
poke$Type.1 <- factor(poke$Type.1, levels = highest_to_lowest)
```
---

## Coordinates and Scales

Standard Plot:
```{r echo = FALSE, out.height= 400, out.width=500}
ggplot(data = poke, mapping = aes(x = Type.1)) +
  geom_bar()
```

x-axis is cluttered!
---
## Coordinates and Scales

Can we switch x for y?
```{r echo = TRUE, out.height= 300, out.width=400}
ggplot(data = poke, mapping = aes(y = Type.1)) +
  geom_bar()
```

Sure! But let's try another methods!
---

## Coordinates and Scales

Coordinate Flip
```{r echo = TRUE, out.height= 300, out.width=400}
ggplot(data = poke, mapping = aes(x = Type.1)) +
  geom_bar()+
  coord_flip()
```
---

## Axis Scaling
Adjusting the axes to represent different scales/information is done through a simple call:

`scale_x_[function]()` or `scale_y_[function]()`

Built in scale functions are: `continuous`, `discrete`, `log10`, `sqrt`, and `reverse`

As we saw last session, `discrete` scale only works for categorical variables (text, factor, integer)

All other scales only work for continuous numeric variables

---

## Axis Scaling

Get some new data

```[r]
mtcars <- mtcars
```
---

## Axis Scaling

Basic Plot; The additional `scale_x_continuous` call is where we will adjust the x-axis

```{r, echo = TRUE, out.height = 200, out.width = 300}
ggplot(data = mtcars)+ 
  geom_point(aes(x = mpg, y = qsec))+
  scale_x_continuous()+ #This is where we can scale the x-axis
  labs(title = "Miles per Gallon vs. Cylinder Count",
       subtitle = "V-Shaped vs. Flat",
       x = "Quarter Mile Time (seconds)",
       y = "Miles per Gallon (mpg)")
```
---
## Axis Scaling

Change the tick marks (`breaks`)

```{r, echo = TRUE, out.height = 200, out.width = 300}
ggplot(data = mtcars)+ 
  geom_point(aes(x = mpg, y = qsec))+
  scale_x_continuous(breaks = c(10,20,30))+ #change the "breaks" (labels)
  labs(title = "Miles per Gallon vs. Cylinder Count",
       subtitle = "V-Shaped vs. Flat",
       x = "Quarter Mile Time (seconds)",
       y = "Miles per Gallon (mpg)")
```
---
## Axis Scaling

Reverse order the tick marks (`breaks`)

```{r, echo = TRUE, out.height = 200, out.width = 300}
ggplot(data = mtcars)+ 
  geom_point(aes(x = mpg, y = qsec))+
  scale_x_reverse()+ 
  labs(title = "Miles per Gallon vs. Cylinder Count",
       subtitle = "V-Shaped vs. Flat",
       x = "Quarter Mile Time (seconds)",
       y = "Miles per Gallon (mpg)")
```
---
## Zoom
Useful for insets
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = mtcars, aes(x = mpg))+
  geom_histogram()
```

Lets focus on cars getting >25mpg
---
## Zoom
using `scale_x_continuous` and the `limits` call we can zoom to a particular section
```{r, echo = TRUE, out.height = 300, out.width = 400, message = FALSE}
ggplot(data = mtcars, aes(x = mpg))+
  geom_histogram()+
  scale_x_continuous(limits = c(25, 35), breaks = seq(25,35,2))
```
---

## Scaling Non-Normal Data

Not all data is well-suited for "normal" distribution plots! Let's load some new data!

```{r}
library(gapminder)
```

---
##Scaling Non-Normal Data

Basic Smooth based on a log-normal distribution
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ log(x))
```

This isn't very attractive! Because our axes are not scaled appropriately!
---

## Scaling Non-Normal Data

Simply adding `scale_log10_x` creates a linear graph showing the trend better
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10()
```
---
## Themes

Themes allow us to change the appearance of ggplots

ggplot contains a number of native themes some better than others

---
## Themes
Theme Minimal:
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type") +
  theme_minimal()
```

---
## Themes
The `theme` call can help us take even more control over the theme like the placement of the title
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type") +
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0.5) # hjust 0 = left, 0.5 = center, 1 = right
  )
```
---

## Themes

How much can we change? Everything!!
```[r]
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type",
       x = "Base Attacking Skills",
       y = "Base Defending Skills") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 20), # adjust the centering and size of title
        legend.position = "bottom", # positions the legend on the bottom
        legend.title = element_text(size = 15), # change the legend title font size
        legend.text = element_text(size = 12), # change the legend label font size
        axis.title.x = element_text(size = 15), # change the axis title font size
        axis.text.x = element_text(size = 15), # change the axis label font size
        axis.title.y = element_text(size = 15), # change the axis title font size
        axis.text.y = element_text(size = 15)) + # change the axis label font size
  scale_color_discrete(name = "Pokemon Type", # make some fancy names in the legend using aesthetic scale
                       labels = c("Fire Pokemon",
                                  "Grass Pokemon",
                                  "Water Pokemon"))
```
---
## Themes

How much can we change? Everything!!
```{r, echo = FALSE}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type",
       x = "Base Attacking Skills",
       y = "Base Defending Skills") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 20), # adjust the centering and size of title
        legend.position = "bottom", # positions the legend on the bottom
        legend.title = element_text(size = 15), # change the legend title font size
        legend.text = element_text(size = 12), # change the legend label font size
        axis.title.x = element_text(size = 15), # change the axis title font size
        axis.text.x = element_text(size = 15), # change the axis label font size
        axis.title.y = element_text(size = 15), # change the axis title font size
        axis.text.y = element_text(size = 15)) + # change the axis label font size
  scale_color_discrete(name = "Pokemon Type", # make some fancy names in the legend using aesthetic scale
                       labels = c("Fire Pokemon",
                                  "Grass Pokemon",
                                  "Water Pokemon"))
```
---
## Themes

This is pretty intense, can you make it easier? 

Use the ESA Template!! Simply copy and paste from lines 260 to 287 to the top of your script

```{r}
library(ggthemr)

esa_colours <- c('#F9A134', '#699CC6', '#A0CF67', '#9370B1', '#F26531', '#72CCD2', '#098A5B', '#781D7E','#B12533', '#004261', '#015E44', '#421953', '#DCDDDE', '#ADAFB2', '#6D6E71')

esa_colours <- c('#000000', esa_colours) #Add black as the "initial" color (outlining boxes, and plain plots)

esa_theme <- define_palette(
  swatch = esa_colours, # colors for plotting points
  gradient = c(lower = esa_colours[7L], upper = esa_colours[11L]), # upper and lower colors for continuous colors
  background = '#FFFFFF' # White background
)
```

---
## Themes

Apply this as a "theme" using `ggthemr`

```{r}
ggthemr(esa_theme, type = "outer", layout = "minimal", spacing = 3)
```
---
## Themes

Build this as a complete theme which can be called in ggplot

```{r}
theme_esa <- function(base_size = 11, base_family = "", base_line_size = base_size/22, base_rect_size = base_size/22){
  theme_bw(base_size = base_size, base_family = base_family,
           base_line_size = base_line_size, base_rect_size = base_rect_size) %+replace%
    theme(panel.border =element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(color = '#000000', size = rel(1)),
          legend.key = element_blank(),
          plot.title = element_text(size = 18, face = 'bold'),
          plot.subtitle = element_text(size = 12, face = 'italic', color = 'grey40'),
          plot.caption = element_text(hjust = 0, size = 9, color = 'grey30'),
          strip.background = element_rect(fill = '#FFFFFF', color = "#000000", size = rel(2)), complete = TRUE)
}
```
---

## Theme

Test our new theme!

```
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type",
       x = "Base Attacking Skills",
       y = "Base Defending Skills") +
  theme_esa() + # THEME IS HERE
   scale_color_manual(name = "Pokemon Type", # make some fancy names in the legend using aesthetic scale
                      labels = c("Fire Pokemon",
                                 "Grass Pokemon",
                                 "Water Pokemon"),
                      values = c(esa_colours[2], esa_colours[4], esa_colours[3])) # scale the colors manually to ensure they match (can we do this another way?)
```
---
Test our new theme!

```{r, echo = FALSE, out.height = 500, out.width = 600}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = Type.1)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type",
       x = "Base Attacking Skills",
       y = "Base Defending Skills") +
  theme_esa() + # THEME IS HERE
   scale_color_manual(name = "Pokemon Type", # make some fancy names in the legend using aesthetic scale
                      labels = c("Fire Pokemon",
                                 "Grass Pokemon",
                                 "Water Pokemon"),
                      values = c(esa_colours[2], esa_colours[4], esa_colours[3])) # scale the colors manually to ensure they match (can we do this another way?)
```
---
