---
title: "ggplot Continued"
author: 
  - "Taylor Spaulding"
date: "2021-02-05"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#699CC6",
  secondary_color = "#F9A134",
  inverse_header_color = "#FFFFFF",
  extra_css = list(".title-slide" = list("background-color" = "#FFFFFF"),
                   ".title-slide, .title-slide h1, .title-slide h2, .title-slide h3" = 
                     list("color" = "#699CC6")))
```

## ggplot Continued   

Topics for Today
  - Axes
  - Theming
  - Intro to gridExtra

---

## Load Up Your Packages

```{r echo = TRUE}
remotes::install_github('Mikata-Project/ggthemr')

library(tidyverse)
library(gridExtra)
library(ggthemr)
```

---

## Axis Scaling
Adjusting the axes to represent different scales/information is done through a simple call:

`scale_x_[function]()` or `scale_y_[function]()`

Built in scale functions are: `continuous`, `discrete`, `log10`, `sqrt`, and `reverse`

As we saw last session, `discrete` scale only works for categorical variables (text, factor, integer)

All other scales only work for continuous numeric variables

---

## Axis Scaling

Get some new data

```[r]
mtcars <- mtcars
```
---

## Axis Scaling

Basic Plot; The additional `scale_x_continuous` call is where we will adjust the x-axis

```{r, echo = TRUE, out.height = 200, out.width = 300}
ggplot(data = mtcars)+ 
  geom_point(aes(x = mpg, y = qsec))+
  scale_x_continuous()+ #This is where we can scale the x-axis
  labs(title = "Miles per Gallon vs. Cylinder Count",
       subtitle = "V-Shaped vs. Flat",
       x = "Quarter Mile Time (seconds)",
       y = "Miles per Gallon (mpg)")
```
---
## Axis Scaling

Change the tick marks (`breaks`)

```{r, echo = TRUE, out.height = 200, out.width = 300}
ggplot(data = mtcars)+ 
  geom_point(aes(x = mpg, y = qsec))+
  scale_x_continuous(breaks = c(10,20,30))+ #change the "breaks" (labels)
  labs(title = "Miles per Gallon vs. Cylinder Count",
       subtitle = "V-Shaped vs. Flat",
       x = "Quarter Mile Time (seconds)",
       y = "Miles per Gallon (mpg)")
```
---
## Axis Scaling

Reverse order the tick marks (`breaks`)

```{r, echo = TRUE, out.height = 200, out.width = 300}
ggplot(data = mtcars)+ 
  geom_point(aes(x = mpg, y = qsec))+
  scale_x_reverse()+ 
  labs(title = "Miles per Gallon vs. Cylinder Count",
       subtitle = "V-Shaped vs. Flat",
       x = "Quarter Mile Time (seconds)",
       y = "Miles per Gallon (mpg)")
```
---
## Zoom
Useful for insets
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = mtcars, aes(x = mpg))+
  geom_histogram()
```

Lets focus on cars getting >25mpg
---
## Zoom
using `scale_x_continuous` and the `limits` call we can zoom to a particular section
```{r, echo = TRUE, out.height = 300, out.width = 400, message = FALSE}
ggplot(data = mtcars, aes(x = mpg))+
  geom_histogram()+
  scale_x_continuous(limits = c(25, 35), breaks = seq(25,35,2))
```
---

## Scaling Non-Normal Data

Not all data is well-suited for "normal" distribution plots! Let's load some new data!

```{r}
library(gapminder)
```

---
##Scaling Non-Normal Data

Basic Smooth based on a log-normal distribution
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ log(x))
```

This isn't very attractive! Because our axes are not scaled appropriately!
---

## Scaling Non-Normal Data

Simply adding `scale_log10_x` creates a linear graph showing the trend better
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10()
```
---
## Themes

Themes allow us to change the appearance of ggplots

ggplot contains a number of native themes some better than others

---

## Load In Some Data

Load in the data from a dataset stored in the Project. 

```{r echo = TRUE}
poke <- read_csv("Pokemon.csv")

poke_mod <- poke %>%
  filter(`Type 1` %in% c("Grass","Water","Fire")) #filter for just those with these types

poke_mod <- poke_mod %>%
  mutate(`Type 1` = factor(`Type 1`, levels = c("Fire","Water","Grass")))

str(poke_mod)

```

---
## Themes
Theme Minimal:
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = `Type 1`)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type") +
  theme_minimal()
```

---
## Themes
The `theme` call can help us take even more control over the theme like the placement of the title
```{r, echo = TRUE, out.height = 300, out.width = 400}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = `Type 1`)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type") +
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0.5) # hjust 0 = left, 0.5 = center, 1 = right
  )
```
---

## Themes

How much can we change? Everything!!
```[r]
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = `Type 1`)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type",
       x = "Base Attacking Skills",
       y = "Base Defending Skills") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 20), # adjust the centering and size of title
        legend.position = "bottom", # positions the legend on the bottom
        legend.title = element_text(size = 15), # change the legend title font size
        legend.text = element_text(size = 12), # change the legend label font size
        axis.title.x = element_text(size = 15), # change the axis title font size
        axis.text.x = element_text(size = 15), # change the axis label font size
        axis.title.y = element_text(size = 15), # change the axis title font size
        axis.text.y = element_text(size = 15)) + # change the axis label font size
  scale_color_discrete(name = "Pokemon Type", # make some fancy names in the legend using aesthetic scale
                       labels = c("Fire Pokemon",
                                  "Grass Pokemon",
                                  "Water Pokemon"))
```
---
## Themes

How much can we change? Everything!!
```{r, echo = FALSE}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = `Type 1`)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type",
       x = "Base Attacking Skills",
       y = "Base Defending Skills") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 20), # adjust the centering and size of title
        legend.position = "bottom", # positions the legend on the bottom
        legend.title = element_text(size = 15), # change the legend title font size
        legend.text = element_text(size = 12), # change the legend label font size
        axis.title.x = element_text(size = 15), # change the axis title font size
        axis.text.x = element_text(size = 15), # change the axis label font size
        axis.title.y = element_text(size = 15), # change the axis title font size
        axis.text.y = element_text(size = 15)) + # change the axis label font size
  scale_color_discrete(name = "Pokemon Type", # make some fancy names in the legend using aesthetic scale
                       labels = c("Fire Pokemon",
                                  "Grass Pokemon",
                                  "Water Pokemon"))
```
---
## Themes

This is pretty intense, can you make it easier? 

Use the ESA Template!! Simply copy and paste from lines 260 to 287 to the top of your script

```{r}
library(ggthemr)

esa_colours <- c('#F9A134', '#699CC6', '#A0CF67', '#9370B1', '#F26531', '#72CCD2', '#098A5B', '#781D7E','#B12533', '#004261', '#015E44', '#421953', '#DCDDDE', '#ADAFB2', '#6D6E71')

esa_colours <- c('#000000', esa_colours) #Add black as the "initial" color (outlining boxes, and plain plots)

esa_theme <- define_palette(
  swatch = esa_colours, # colors for plotting points
  gradient = c(lower = esa_colours[7L], upper = esa_colours[11L]), # upper and lower colors for continuous colors
  background = '#FFFFFF' # White background
)
```

---
## Themes

Apply this as a "theme" using `ggthemr`

```{r}
ggthemr(esa_theme, type = "outer", layout = "minimal", spacing = 3)
```
---
## Themes

Build this as a complete theme which can be called in ggplot

```{r}
theme_esa <- function(base_size = 11, base_family = "", base_line_size = base_size/22, base_rect_size = base_size/22){
  theme_bw(base_size = base_size, base_family = base_family,
           base_line_size = base_line_size, base_rect_size = base_rect_size) %+replace%
    theme(panel.border =element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(color = '#000000', size = rel(1)),
          legend.key = element_blank(),
          plot.title = element_text(size = 18, face = 'bold'),
          plot.subtitle = element_text(size = 12, face = 'italic', color = 'grey40'),
          plot.caption = element_text(hjust = 0, size = 9, color = 'grey30'),
          strip.background = element_rect(fill = '#FFFFFF', color = "#000000", size = rel(2)), complete = TRUE)
}
```
---

## Theme

Test our new theme!

```
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = `Type 1`)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type",
       x = "Base Attacking Skills",
       y = "Base Defending Skills") +
  theme_esa() + # THEME IS HERE
   scale_color_manual(name = "Pokemon Type", # make some fancy names in the legend using aesthetic scale
                      labels = c("Fire Pokemon",
                                 "Grass Pokemon",
                                 "Water Pokemon"),
                      values = c(esa_colours[2], esa_colours[4], esa_colours[3])) # scale the colors manually to ensure they match (can we do this another way?)
```
---
Test our new theme!

```{r, echo = FALSE, out.height = 500, out.width = 600}
ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense, color = `Type 1`)) +
  geom_point() +
  labs(title = "Defending/Attacking Scatter Plot By Pokemon Type",
       x = "Base Attacking Skills",
       y = "Base Defending Skills") +
  theme_esa() + # THEME IS HERE
   scale_color_manual(name = "Pokemon Type", # make some fancy names in the legend using aesthetic scale
                      labels = c("Fire Pokemon",
                                 "Grass Pokemon",
                                 "Water Pokemon"),
                      values = c(esa_colours[2], esa_colours[4], esa_colours[3])) # scale the colors manually to ensure they match (can we do this another way?)
```

---

Remove `ggthemr` which does not function well with gridExtra anymore

```{r, eval = TRUE, echo = TRUE}

detach("package:ggthemr", unload = TRUE)

```

---