---
title: "ggplot Continued"
author: 
  - "Taylor Spaulding"
date: "2021-01-22"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#699CC6",
  secondary_color = "#F9A134",
  inverse_header_color = "#FFFFFF",
  extra_css = list(".title-slide" = list("background-color" = "#FFFFFF"),
                   ".title-slide, .title-slide h1, .title-slide h2, .title-slide h3" = 
                     list("color" = "#699CC6")))
```

## ggplot Continued   

Topics for Today
  - Intro to gridExtra

---

## Load Up Your Packages

```{r echo = TRUE}
library(tidyverse)
library(gridExtra)
```
---

## Load In Some Data

Load in the data from a dataset stored in the Project. 

```{r echo = TRUE}
poke <- read_csv("C:/Users/tspaulding/Documents/R/R-Intro-Training/Data/Pokemon.csv")

poke_mod <- poke %>%
  filter(`Type 1` %in% c("Grass","Water","Fire")) #filter for just those with these types

poke_mod <- poke_mod %>%
  mutate(`Type 1` = factor(`Type 1`, levels = c("Fire","Water","Grass")))

str(poke_mod)

```

---
## Basic Plots

Build a few Basic plots which we can maneuver

``` {r echo = TRUE}
atk_def <- ggplot(data = poke_mod, mapping = aes(x = Attack, y = Defense)) +
  geom_point(aes(color = `Type 1`)) + theme(legend.position="bottom")

atk_hp <- ggplot(data = poke_mod, mapping = aes(x = Attack, y = HP)) +
  geom_point(aes(color = `Type 1`)) + theme(legend.position="bottom")

spd_hp <- ggplot(data = poke_mod, mapping = aes(x = Speed, y = HP)) +
  geom_point(aes(color = `Type 1`)) + theme(legend.position="bottom")

spatk_spdef <- ggplot(data = poke_mod, mapping = aes(x = `Sp. Atk`, y = `Sp. Def`))+ 
  geom_point(aes(color = `Type 1`)) + theme(legend.position="bottom")
```

---
## gridArrange

Let's arrange the four plots into a 2x2 grid

```{r}
grid.arrange(atk_def, atk_hp, spd_hp, spatk_spdef, ncol = 2, nrow = 2)
```

---
## Legend Management

Four legends are crowding the plot if we have a common legend, we can extract the legend with a custom fuction (This will work for any plot)
```{r echo = TRUE}
g_legend<-function(plot){
  tmp <- ggplot_gtable(ggplot_build(plot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
```

---
## Legend Management

Use the function on a single plot to extract the legend

```{r echo = TRUE}
mylegend<-g_legend(atk_def)
```

---
## Legend Management

We can use a rather complicated arrangement to place the legend underneath the 2x2 matrix


```
grid.arrange(arrangeGrob(atk_def + theme(legend.position = "none"),
             atk_hp + theme(legend.position = "none"),              
             spd_hp + theme(legend.position = "none"),              
             spatk_spdef + theme(legend.position = "none"),         
             ncol = 2, nrow = 2),                                   
             mylegend, nrow = 2, heights = c(30, 2))
    
```

The nested `arrangeGrob()` allows for us to create the 2x2 matrix as a single grob (graphic object).

In the main `grid.arrange()`, the first grob is the matrix and the second is the legend we created 


---
## Legend Management

``` {r echo = FALSE}
grid.arrange(arrangeGrob(atk_def + theme(legend.position = "none"),
             atk_hp + theme(legend.position = "none"),              
             spd_hp + theme(legend.position = "none"),              
             spatk_spdef + theme(legend.position = "none"),         
             ncol = 2, nrow = 2),                                   
             mylegend, nrow = 2, heights = c(30, 2))
```
---
## gridExtra and Beyond

Can make very complicated plots by using custom layouts

```{r}
layout = rbind(c(1,1,1,1,1,NA),
               c(2,2,2,2,2,3),
               c(2,2,2,2,2,3),
               c(2,2,2,2,2,3),
               c(2,2,2,2,2,3),
               c(2,2,2,2,2,3))
```
---
## gridExtra and Beyond

make a histogram of attack scores...

```{r}
atk_histo <- ggplot(data = poke_mod, mapping = aes(x = Attack, fill = `Type 1`))+
  geom_histogram()+
  theme(
    legend.position = "bottom",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )
```
---
## gridExtra and Beyond

...and of defence scores (notice the coord_flip to make it vertical)

```{r}
def_histo <- ggplot(data = poke_mod, mapping = aes(x = Defense, fill = `Type 1`))+
  geom_histogram()+
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )+
  coord_flip()
```
extract the legend
```{r}
mylegend<-g_legend(atk_histo)
```
---
## gridExtra and Beyond

Plot it

```{r eval = FALSE}
grid.arrange(arrangeGrob(atk_histo + theme(legend.position = "none"), 
                         atk_def + theme(legend.position = "none") + labs(x = "Attack"), 
                         def_histo, 
                         layout_matrix = layout),
             mylegend, 
             nrow = 2, heights = c(30,2))
```
---
## gridExtra and Beyond

Plot it

```{r echo = FALSE, message = FALSE}
grid.arrange(arrangeGrob(atk_histo + theme(legend.position = "none"), 
                         atk_def + theme(legend.position = "none") + labs(x = "Attack"), 
                         def_histo, 
                         layout_matrix = layout),
             mylegend, 
             nrow = 2, heights = c(30,2))
```

