---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(esaRmisc)
knitr::opts_chunk$set(echo = FALSE)
```


## R Basics

### R and RStudio

```{r r-rstudio}
quiz(
  question("You can use R without installing RStudio",
    answer("True", correct = TRUE),
    answer("False")
  ),
  question("You can use RStudio without installing R",
    answer("True"),
    answer("False", correct = TRUE)
  )
)
```

### Packages

```{r packages}
quiz(
  question("What is the main package repository for R?",
    answer("CPAN"),
    answer("CRAN", correct = TRUE),
    answer("PyPI"),
    answer("Cargo"),
    answer("NuGet")
  ),
  question("Which package allows you to install packages directly from GitHub?",
    answer("BiocManager"),
    answer("githubR"),
    answer("remotes", correct = TRUE),
    answer("installr")
  )
)
```

### Vectors 

Fix the following code to create a vector with the numbers 1, 2, 3:

```{r vec-123, exercise=TRUE}
(1 2 3)
```

The following code creates a vector of the whole numbers from 1 to 5. Write code that creates a vector with whole numbers from 5 to 1. 
```{r seq-rev, exercise=TRUE, exercise.eval=TRUE}
1:5
```

```{r seq-rev-hint}
5:1
seq(5, 1, -1)
rev(1:5)
```

Fix the code to create a vector that ranges from 3 to 1 in increments of 0.5.
```{r seq-err, exercise=TRUE}
seq(1, 3, -0.5)
```

```{r seq-err-hint}
seq(3, 1, -0.5)
seq(to = 1, from = 3, by = -0.5)
```

Add a line of code to multiply every element of the vector `x` by 2. 
```{r div-2, exercise=TRUE, exercise.eval=TRUE}
x <- seq(1, 5, 0.5)
```

```{r div-2-hint}
x*2
```

Fix the error in this code.
```{r miss-quotes, exercise=TRUE}
paste0(c("flood", "dragon"), c(plain, "fly"))
```

```{r miss-quotes-hint}
paste0(c("flood", "dragon"), c("plain", "fly"))
```

### Tibble

Fix the errors in this code.
```{r tibble-err, exercise=TRUE}
tibble(Integer = 1L:3L
       Double = c(1, 2.5; 3),
       Logical = c(TRUE FALSE TRUE),
       Character = ("strings", "better", "name"),
       Factor = as.factor(c("factors", "are", "special")))
```

```{r tibble-err-hint}
tibble(Integer = 1L:3L,
       Double = c(1, 2.5, 3),
       Logical = c(TRUE, FALSE, TRUE),
       Character = c("strings", "better", "name"),
       Factor = as.factor(c("factors", "are", "special")))
```

## Data Transformation

### Filter

Change the code to return rows where `WaterYearType` is `Wet`.
```{r filter-wyt, exercise=TRUE}
filter(water_year_type, WaterYearType == "Above Normal")
```

Write code to return rows from `water_year_type` where `WaterYearType` is `Critical` in the `SAC` valley.
```{r filter-crit, exercise=TRUE}

```

```{r filter-crit-hint}
filter(water_year_type, WaterYearType == "Critical" & Valley == "SAC")
```

Write code to return rows from `water_year_type` where `WaterYearType` is `NA`.
```{r filter-na, exercise=TRUE}

```

```{r filter-na-hint}
filter(water_year_type, is.na(WaterYearType))
```

Write code to return rows from `water_year_type` where `WaterYearType` is not `NA`.
```{r filter-not-na, exercise=TRUE}

```

```{r filter-not-na-hint}
filter(water_year_type, !is.na(WaterYearType))
```

Write code to return rows from `water_year_type` where `WaterYearType` is `NA` or `Dry`.
```{r filter-na-or, exercise=TRUE}

```

```{r filter-na-or-hint}
filter(water_year_type, is.na(WaterYearType) | WaterYearType == "Dry")
```

Write code to return rows from `water_year_type` where `WaterYearType` is `Critical` or `Dry`.
```{r filter-in, exercise=TRUE}

```

```{r filter-in-hint}
filter(water_year_type, WaterYearType %in% c("Critical", "Dry"))
# alternatively (but less concise)
filter(water_year_type, WaterYearType == "Critical" | WaterYearType == "Dry")
```

Write code to return rows from `water_year_type` where `WaterYearType` is not `Critical` or `Dry`.
```{r filter-not-in, exercise=TRUE}

```

```{r filter-not-in-hint}
filter(water_year_type, !(WaterYearType %in% c("Critical", "Dry")))
# note that NAs are retained
```

Write code to return rows from `water_year_type` where `WaterYear` is 2001-2010.
```{r filter-years, exercise=TRUE}

```

```{r filter-years-hint}
filter(water_year_type, WaterYear %in% 2001:2010)
filter(water_year_type, WaterYear >= 2001 & WaterYear <= 2010)
filter(water_year_type, WaterYear > 2000 & WaterYear < 2011)
```

### Arrange

Change code to arrange rows from `water_year_type` by descending `WaterYear`.
```{r arrange-desc, exercise=TRUE}
arrange(water_year_type, WaterYear)
```

```{r arrange-desc-hint}
arrange(water_year_type, desc(WaterYear))
```

Write code to arrange rows from `water_year_type` by `WaterYearYearType`, descending `Valley`, and `WaterYear`. Note, sort order for factors is special. We will come back to this later in the tutorial. 
```{r arrange-multi, exercise=TRUE}

```

```{r arrange-multi-hint}
arrange(water_year_type, WaterYearType, desc(Valley), WaterYear)
```

### Select

Change code to select columns from `water_year_type` such that the `Valley` column is on the left.
```{r select-order, exercise=TRUE}
select(water_year_type, WaterYear, Valley, WaterYearType)
```

```{r select-order-hint}
select(water_year_type, Valley, WaterYear, WaterYearType)
```

Write code to drop `WYT` column from `water_year_type`.
```{r select-drop, exercise=TRUE}

```

```{r select-drop-hint}
select(water_year_type, -WYT)
```

Write code to drop `WYT` column from `water_year_type` and rename `WaterYearType` to `WYT`.
```{r select-drop-rename, exercise=TRUE}

```

```{r select-drop-rename-hint}
rename(select(water_year_type, -WYT), WYT = WaterYearType)
select(water_year_type, WaterYear, Valley, WYT = WaterYearType)
```

### Rename

Write code to rename `WYT` to `WYTCode` and `WaterYearType` to `WYT`.
```{r rename, exercise=TRUE}

```

```{r rename-hint}
rename(water_year_type, WYTCode = WYT, WYT = WaterYearType)
```

### Mutate

Before we get into mutate, let's revisit what I mentioned earlier about sorting of factors. Characters are sorted alphabetically. Factors are sorted according to a specified order, which defaults to alphabetical. In the case of `water_year_type`, the order of the `WYT` (and `WaterYearType`) levels was previously specified to sort from driest to wettest. In contrast, `Valley` is a character vector that sorts in alphabetical order.

```{r, echo=TRUE}
sort(unique(water_year_type$WYT))
sort(unique(water_year_type$Valley))
```

Use the function `as.character` to create a new column `WYT_char` that converts `WYT` to a character vector.
```{r mutate, exercise=TRUE}

```

```{r mutate-hint}
mutate(water_year_type, WYT_char = as.character(WYT))
```

Change the 2nd line of code to create a new factor column, `WYT_fac`, with levels arranged from wettest to driest. 
```{r mutate-fac, exercise=TRUE}
water_year_type2 <- mutate(water_year_type, WYT_char = as.character(WYT))
mutate(water_year_type2, WYT_fac = factor(WYT_char, levels = c("AN", "BN", "C", "D", "W")))
```

```{r mutate-fac-hint}
water_year_type2 <- mutate(water_year_type, WYT_char = as.character(WYT))
mutate(water_year_type2, WYT_fac = factor(WYT_char, levels = c("W", "AN", "BN", "D", "C")))
```

### Arrange Revisited

Change code to see difference between arranging rows based on `WYT`, `WYT_char`, and `WYT_fac`.
```{r arrange-revisited, exercise=TRUE, exercise.eval=TRUE}
water_year_type %>% 
  mutate(WYT_char = as.character(WYT),
         WYT_fac = factor(WYT_char, levels = c("W", "AN", "BN", "D", "C"))) %>% 
  arrange(WYT)
```

### Group by and summarise

Change code to achieve same result but use `length` instead of `n`.
```{r group-n, exercise=TRUE}
water_year_type %>% 
  group_by(WaterYearType) %>% 
  summarise(Count = n())
```

```{r group-n-hint}
water_year_type %>% 
  group_by(WaterYearType) %>% 
  summarise(Count = length(WaterYear))
```

Change code to achieve same result with `count`.
```{r count, exercise=TRUE}
water_year_type %>% 
  group_by(WaterYearType) %>% 
  summarise(Count = n())
```

```{r count-hint}
water_year_type %>% 
  count(WaterYearType)
```

Write code to `count` by `Valley` and `WaterYearType`.
```{r count2, exercise=TRUE}

```

```{r count2-hint}
water_year_type %>% 
  count(Valley, WaterYearType)
```

Write code to remove `NA` in `WaterYearType` before using `count` by `Valley` and `WaterYearType`.
```{r count3, exercise=TRUE}

```

```{r count3-hint}
water_year_type %>% 
  filter(!is.na(WaterYearType)) %>% 
  count(Valley, WaterYearType)
```

Write code to remove `NA` in `WaterYearType`, `count` by `Valley` and `WaterYearType`, and then calculate the proportion of years in each water year type.
```{r count4, exercise=TRUE}

```

```{r count4-hint}
water_year_type %>% 
  filter(!is.na(WaterYearType)) %>% 
  count(Valley, WaterYearType) %>% 
  group_by(Valley) %>% 
  mutate(PropYears = n/sum(n))
```